---
title: Approximating the global economic (market) value of farmed animals
output: 
      md_document:
        variant: markdown_github
        toc: true 
html_preview: false
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```

```{r echo=FALSE, results="hide", message=FALSE}
suppressPackageStartupMessages({
  library(badger)
  library(here)
  library(knitr)
})
``` 

```{r, echo = FALSE, results='asis'}
cat(
  badge_license("CC BY-SA 4.0"),
  badge_lifecycle("experimental")
)
```



# Description
 
This repository contains all `R` code, output data and figures for the 
[*GBADS*](https://animalhealthmetrics.org/) paper 
"Approximating the Global Economic (Market) Value of Farmed Animals".

>> Schrobback, Peggy, Gabriel Dennis, Yin Li, Dianne Mayberry, Alexandra
>> Shaw, Theodore Knight-Jones, Thomas Lloyd Marsh, et al. 2023.
>> “Approximating the Global Economic (Market) Value of Farmed Animals.”
>> *Global Food Security* 39: 100722.
>> <https://doi.org/https://doi.org/10.1016/j.gfs.2023.100722>.
 

# Installation 

To install this package run the following code 

```{r eval=FALSE}
if (!require(remotes)) {
  install.packages("remotes")
}

remotes::install_github(repo = "GBADsInformatics/TEVProject")
```

# Repository Structure 

This repository is structured similar to an R package, however, there are
minimal differences which make this package non-conforming to certain standards.

## Make

The `R` directory contains any common functions which are used across the entire
project. The `inst` directory contains `R` scripts which are intended to be run
via the GNU `Makefile` which specifies the required processes to build each
target. Targets are run using the syntax

```{bash, eval = FALSE}
make name-of-file-to-make
```

Currently, not all the dependencies for the targets which download data are
specified in subsequent targets, due to some side effects of the download
process, however, all files in the `data/processed` and `data/output`
directories should be targets which can be recreated by the project `Makefile`. 


Project locations and directory structure are outlined in the project
configuration file `conf/config.yml`, in accessing these parameters in code is
managed by the `R` package `config`. The location of the configuration file can
be modified, however, if this is done, then one must also modify the environment
variable `R_CONFIG_FILE` to match this new location.
 
Environment and package management is done using `renv`, `renv.lock` specifies
the hashes for each package used.  Package imports are also listed in the
package `DESCRIPTION` file.

## Data 


Currently, the processed and output data is located in `data/processed/` and
`data/output`. Source data is located in `data/source` and is not included in
this repository due to its size.  The versions of the source data which was downloaded for
this project is backed up on an internal *CSIRO* cloud storage platform.


Output data is primarily stored in `parquet` files in the `data/output`
directory.

*(Note: These locations are specified in the project configuration yaml file)*

### Metadata 

Metadata for this project in JSON format for the GBADS Knowledge Engine 
is stored in the directory `data/metadata` and can be 
generated using the Makefile target `make data/metadata`. 


## Figures and Tables 

Output figures and tables are located in `output/figures/`, `output/tables/`. 

*(Note: This is specified in the project configuration yaml file)*

## Codes 

FAOSTAT item codes and country codes are located in `data/codes`. 

*(Note: This is specified in the project configuration yaml file)*

##  Documentation

[pkgdown](https://pkgdown.r-lib.org/) Documentation is available in the 
`docs/` directory and can be accessed locally.

## Figures 

Paper figures can be seen in the [figures](output/figures/README.md) directory.
 
# Licence 
 
This project is under a creative commons licence.


