---
title: "GBADS Livestock Value Estimation Supplementary Notes"
author: PPSTheme
output: 
    pdf_document:
        toc: true
        df_print: kable
vignette: >
  %\VignetteIndexEntry{Documentation}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
  
bibliography: ../references.bib
csl: "../cambridge-university-press-numeric.csl"
link-citations: true
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  echo = FALSE,
  comment = "#>"
)

knitr::opts_knit$set(
  root.dir = here::here(".")
)

suppressPackageStartupMessages({
  library(DiagrammeR) # For the flow diagrams
})
```


```{r config}
config <- config::get()
```


```{r print-table}
print_table <- function(data, caption) {
  data |>
    knitr::kable(caption = caption) |>
    kableExtra::kable_styling(font_size = 6)
}
```

\newpage



# Calculations 


The following section details the process of obtaining both asset value and 
output value estimates for terrestrial livestock and aquaculture. 

### Livestock 



\newpage

## Asset Value Calculation


The following set of equations outlines the terms and calculations used to
obtain the livestock asset value for a single livestock animal type in a single
country for one year during the observational period.


$$
\begin{aligned}
    \text{Carcass Yield} &= \text{No. Slaughtered (Head)} \times \text{No. Slaughtered (Tonnes)} \\
    \text{Average Live Body Weight}&= \text{Carcass Yield} \times \text{Carcass Percent}^{-1} \\
    \text{Asset Value} &= \text{Price per Tonne} \times \text{Average Live Bodyweight (tonnes)} \times \text{Livestock Population}
\end{aligned}
$$

The following flow chart describes the process of calculating the asset value 
for a single livestock animal type in a single country for one year 
during the observational period. 



```{r, fig.cap="Livestock Asset Value Calculation Flow Diagram"}
mermaid("graph TB;
        A[Number Livestock Slaughtered]-->C;
        B[Quanitity Livestock Slaughtered]-->C;
        C[Tonnes/Head = Carcass Yield] -->E;
        D[Carcass Percentage]-->E;
        E[Yield/Carcass Percentage = Average Live Body Weight]-->G;
        F[Liveweight Price Per Tonne]-->G;
        G[Liveweight Price Per Tonne/Average Live Body Weight = Price Per Live Animal]-->I;
        H[Livestock Population]-->I;
        I[Asset Value];
        ")
```

\newpage

## Constant $2014-2016$ USD Calculations 

To convert currencies into constant $2014-2016$ USD the following procedure is used. 

1. Obtain the weighted mean price for each commodity in Local Currency units
for the base period of $2014-2016$. Where the weights applied to each of these
prices are the annual production value when being used to calculate livestock or
Aquaculture outputs in constant USD, or a countries amount of stock when being
used to calculate asset values. In the case where a price is missing for one of
these three years, the remaining available years are used without imputation of
an additional value.

2. Determine the weighted average annual exchange rate to *USD* from *LCU*  
using the same weighting scheme as mentioned in the previous point. 

3. Convert the $2014-2016$ Weighted average prices in Local Currency Units to 
*USD* using the weighted average exchange rate derived in the previous step. 

The flow of this calculation is outlined in the following figure  

```{r constant-usd-figure, fig.cap="Constant 2014-2016 USD Calculation Flow Diagram"}
mermaid("graph LR;
        A[2014-2016 Commodity Producer Price  in LCU]-->C;
        B[Livestock Commodity Quantity/Livestock Head]-->C;
        C[Weighted mean 2014-2016 Producer Price in SLC] -->E;
        D[LCU to USD 2014-2016 Average Annual Exchange Rate] -->G;
        F[Livestock Commodity Quantity/Livestock Head]-->G;
        G[Weighted Average Annual LCU to USD 2014-2016 Exchange Rate]-->E;
        E[Constant 2014-2016 USD Value per Tonne]
")
```


Or to show this in a more mathematical manner, let
$\mathbb{Y} = \{2014, 2015, 2016\}$ be the base years used in the constant dollars 
calculation.

If we then let 

$$
x_i  \qquad i \in \mathbb{Y}
$$

be the number of livestock or quantity of livestock output produced in the base
period and $P_i,   \ i \in \mathbb{Y}$ be the *Producer Price* in standard local
currency in that year and $E_i \ i \in \mathbb{Y}$ be the mean average local
currency unit to USD  exchange rate as reported by the World Bank *PA.NUS.ATLS*
or *PA.NUS.FCRF*.

The value in constant *USD* is then 

$$
\left(\sum_{i \in \mathbb{Y}} P_ix_i \right) \left(\sum_{i \in \mathbb{Y}}x_iE_i\right) \qquad = \qquad \text{Constant 2014-2016 USD Value Per Tonne}
$$


This procedure was used to match the process used by the FAO in creating the
Producer Price Index for each commodity item. A link to the document in which
this procedure in described is available
[here](https://fenixservices.fao.org/faostat/static/documents/PP/PP_e.pdf)
(Valid Link as of 25-07-2022).

## Uncertainty 

As is stated in the following datasets metadata files on the main 
sources of error. 

* [FAOSTAT QCL](https://www.fao.org/faostat/en/#data/QCL/metadata)
* [FAOSTAT PP](https://www.fao.org/faostat/en/#data/PP/metadata)
* [FAOSTAT QV](https://www.fao.org/faostat/en/#data/QV/metadata)


> "It is not possible to assess the overall accuracy of the dataset, as the
> source data is largely collected by member countries. For data that is imputed
> (including for non-reporting countries or for products for which countries do
> not collect official data), the accuracy of the imputations is unknown. However,
> imputation processes are assessed and adjusted on a regular basis."
>
> Overall Accuracy (QCL, QV) (2022-08-08)



> "No information is available regarding sampling error. In the EU, regulations
> mandate that the coefficient of variation shall not exceed 3% for the area of
> cultivation for main crops. For non-EU countries, the coefficient of variation
> may be significantly larger. For further information, see country metadata (if
> available)."
> 
> Sampling Error (QCL, QV) (2022-08-08)


> "No information is available regarding the magnitude of non-sampling errors.
> Non-sampling errors may occur as a result of measurement errors, caused by a
> lack of harmonization in statistical methods. For instance, significant
> measurement errors may be present in instances where FAO concepts do not fit
> with national concepts."
> 
> Non-sampling error  (QCL, QV) (2022-08-08)

This makes any calculation of the uncertainty in these estimates difficult, due
to the poor and inconsistent quality of the underlying datasets due to
differences in countries reporting periods, methodologies and interpretation of
different commodities classifications.

\newpage
 
# Codes 

## Aquaculture

The main unit that is used to subset the data is the custom grouping 

> "Fish, crustaceans and molluscs, etc.."

which is what is used in the FAOSTAT yearbook. This subsets the species used to
the main commercial products and excluded production figures for aquatic plants,
pearls and mother-of-pearl. This subsetting and grouping  is outlined in the
Global Aquaculture Production database metadata at `Aqua_E.html`, other notes on
the data are also given in `NOTES_AQUACULTURE_COUNTRY_EN.html`[@FAO2021b].
 
\newpage
 
 
## Crops 

Crop mappings from 

The following table shows the crop codes used [@FAO2021c]. 

```{r}
readRDS(here::here("data", "codes", "faostat", "FAOSTAT_CPC_cropItemCodes.Rds")) |>
  print_table(caption = "FAOSTAT to CPC Crop Codes and Groupings")
```

\newpage





## Value of Production

The following table shows the FAOSTAT Value of Agricultural Production [@FAO2021] 
item codes used. 

- For this paper project only the `Non-Indigenous` value of production items 
were used. 


```{r}
readRDS(
  here::here(config$data$codes$faostat$dir, "FAOSTAT_VOP_Items_NonIndigenous_Codes.rds")
) |>
  print_table(caption = "FAOSTAT Value of Production Livestock Output Item Codes Used")
```


\newpage

## Producer Prices Meat Live Weight

The following table shows the "Meat, live weight <animal>" producer price 
codes used in the FAOSTAT database. 


```{r}
readRDS(here::here(config$data$codes$faostat$dir, "FAOSTAT_Meat_Live_Weight_Codes.rds")) |>
  dplyr::distinct(`Item Code`, .keep_all = TRUE) |>
  print_table(caption = "FAOSTAT Producer Prices Live Weight Item Codes used")
```

\newpage


## FAOSTAT Livestock Items 

The following table displays the Livestock Species item codes which are 
used in this analysis. These codes are used to produce the asset value for each 
species. The number of species may of course differ for each country. 

```{r}
readRDS(
  here::here(config$data$codes$faostat$dir, "FAOSTAT_Livestock_Codes.rds")
) |>
  dplyr::filter(`Domain Code` == "QCL") |>
  dplyr::distinct(Item, .keep_all = TRUE) |>
  print_table(
    caption = "FAOSTAT Production and Livestock Products Livestock (Stock) Item Codes Used"
  )
```




## Country ISO3 Codes 

The following table shows the ISO3 codes for regions which are included in this
analysis, including the years in which the code designation applies.


```{r}

arrow::read_parquet(here::here(config$data$codes$faostat$country_codes$output_codes)) |>
  dplyr::select(country, m49_code, iso3_code, start_year, end_year) |>
  print_table(caption = "ISO3 Country Codes Included in this Analysis")
```


These codes are sourced directly from the following location 

```{r, results='asis'}
config$data$codes$faostat$country_codes$url
```


\newpage

## Technical Conversion Factors 

Technical conversion factors for animal body weights are derived directly from
FAOSTAT production data, or using the values given in the in their technical
conversion factors documentation [@faoTechnicalConversionFactors2012].


*Note: There are inconsistencies in how different countries report body weights,
with some reporting only sold animals, and others the average live body weight
of their entire stock. The carcass percentage also varies per animal depending
on what value is included in the output FAOSTAT Commodity tree.*

In the cases where countries have missing values for different livestock types
these values have been imputed using the regional median value for each of the
live body weight, carcass weight and carcass percentage.

The following table shows these a glimpse at these values once they are 


```{r}
arrow::read_parquet(
  here::here(config$data$output$technical_conversion_factors_imputed)
) |>
  dplyr::glimpse()
```



# References
