---
title: "Frequently Asked Questions"
output:
  html_document:
    theme: cerulean
    toc: yes
    toc_depth: 3
    toc_float: yes
    fig_height: 6
    fig_width: 10
    fig_retina: 3
    highlight: pygments
    highlight_downlit: yes
    code_folding: hide
    number_sections: no
    df_print: paged
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
knitr::opts_knit$set(
  root.dir = here::here(".")
)
```

```{r setup, include=FALSE}

suppressPackageStartupMessages({
  library(arrow)
  library(dplyr)
  library(tidyr)
  library(LivestockValueGBADS)
})


# Project configuration
config <- config::get()
```


# Introduction 

This vignette meant to hold answers to any frequently asked questions. 


# Does animal live body weight exceed output mass? 

To answer this question we can break down the question into the following
three different sub-questions. 


## Does animal live body weight exceed output mass in the cases where we have a price value as a whole?

*Specifically a price value in Constant 2014-2016 USD.*



```{r}

# Load in the livestock output files  and aggregate to
# Live Body Weight
livestock_value <- arrow::read_parquet(config$data$output$livestock_values)


livestock_value |>
  dplyr::select(
    iso3_code,
    item,
    animal,
    year,
    tonnes,
    gross_production_value_constant_2014_2016_thousand_us,
    stock_value_constant_2014_2016_usd
  ) |>
  dplyr::filter(
    # Filter observations in which a year is present
    (item == "stock" & stock_value_constant_2014_2016_usd > 0) |
      (item != "stock" & gross_production_value_constant_2014_2016_thousand_us > 0)
  ) |>
  dplyr::mutate(
    item = ifelse(item == "stock", "live", "output")
  ) |>
  dplyr::group_by(
    item, year
  ) |>
  dplyr::summarise(
    tonnes = sum(tonnes, na.rm = TRUE)
  ) |>
  tidyr::spread(item, tonnes) |>
  dplyr::mutate(
    live_body_weight_exceeds_output = live > output,
    ratio = live / output
  ) |>
  knitr::kable(
    caption = "Does animal live body weight exceed output mass in the cases where we have a price value as a whole?",
    col.names = c(
      "Year",
      "Live Animal Body Mass (Tonnes)",
      "Animal Output Mass (Tonnes)",
      "Does Body Mass Exceed Output Mass?",
      "Live Mass to Output Mass Ratio"
    )
  )
```

For items in which we have a currently assigned value,
output mass is approximately 2 times the live body weight mass. 

*Note: this is partially due to missing values for live weight prices*

## Does animal live body weight  exceed output mass in total? 

*Note: in this instance we are comparing all live body weight and output values, 
regardless if a value has been able to be attached to them.*

```{r}


livestock_value |>
  dplyr::select(
    iso3_code,
    item,
    animal,
    year,
    tonnes,
    gross_production_value_constant_2014_2016_thousand_us,
    stock_value_constant_2014_2016_usd
  ) |>
  dplyr::mutate(
    item = ifelse(item == "stock", "live", "output")
  ) |>
  dplyr::group_by(
    item, year
  ) |>
  dplyr::summarise(
    tonnes = sum(tonnes, na.rm = TRUE),
    .groups = "drop"
  ) |>
  tidyr::drop_na(item) |>
  tidyr::spread(item, tonnes) |>
  dplyr::mutate(
    live_body_weight_exceeds_output = live > output,
    ratio = live / output
  ) |>
  knitr::kable(
    caption = "Does animal live body weight exceed output mass in total?",
    col.names = c(
      "Year",
      "Live Animal Body Mass (Tonnes)",
      "Animal Output Mass (Tonnes)",
      "Does Body Mass Exceed Output Mass?",
      "Live Mass to Output Mass Ratio"
    )
  )
```

For the total value of output mass to live weight mass (with and without values), 
the ratio of live mass to output mass goes from 0.9 in 1994, to 0.7 in 2018. 


## Does animal live body weight exceed output mass for items in which we have a value for both items (within country and year)?

*Note: in this instance we are comparing items in which we both have a stock value, 
and output values for that particular animal.*

```{r}

livestock_value |>
  dplyr::select(
    iso3_code,
    item,
    animal,
    year,
    tonnes,
    gross_production_value_constant_2014_2016_thousand_us,
    stock_value_constant_2014_2016_usd
  ) |>
  tidyr::drop_na(tonnes) |>
  dplyr::filter(
    # Filter observations in which a year is present
    (item == "stock" & stock_value_constant_2014_2016_usd > 0) |
      (item != "stock" & gross_production_value_constant_2014_2016_thousand_us > 0)
  ) |>
  dplyr::mutate(
    value = ifelse(item == "stock", stock_value_constant_2014_2016_usd, gross_production_value_constant_2014_2016_thousand_us)
  ) |>
  dplyr::select(
    iso3_code, item, animal, year, tonnes, value
  ) |>
  dplyr::mutate(
    item = ifelse(item == "stock", "live", "output")
  ) |>
  dplyr::group_by(iso3_code, year, item, animal) |>
  dplyr::summarise_all(~ sum(.x, na.rm = TRUE)) |>
  dplyr::ungroup() |>
  dplyr::group_by(iso3_code, year, animal) |>
  dplyr::mutate(
    g_size = n()
  ) |>
  dplyr::ungroup() |>
  dplyr::filter(g_size == 2) |>
  dplyr::group_by(year, item) |>
  dplyr::summarise(
    tonnes = sum(tonnes, na.rm = TRUE),
    .groups = "drop"
  ) |>
  tidyr::spread(item, tonnes) |>
  dplyr::mutate(
    live_body_weight_exceeds_output = live > output,
    ratio = live / output
  ) |>
  knitr::kable(
    caption = "Does animal live body weight exceed output mass for items in which we have a value for both items?",
    col.names = c(
      "Year",
      "Live Animal Body Mass (Tonnes)",
      "Animal Output Mass (Tonnes)",
      "Does Body Mass Exceed Output Mass?",
      "Live Mass to Output Mass Ratio"
    )
  )
```

For the items in which there is a value for both items, output mass is always
greater than live animal body mass. The ratio of live mass to output mass goes
from 0.86 in 1994, to 0.77 in 2018.



## Caveats

*Note: that as this uses live body weight at slaughter, it is an upper bound on
the average live body weight,  as it is not adjusted for the age and breed
structure present in each country.*

*Note: the above results do not include Aquaculture.*
